Parallel multi-objective optimization using co-evolutionary islands with overlapping search space

### 2016-12-18 hpmoon-coor.tex
A few changes and suggestions here and there

Evolutionary algorithms (EAs) \citep{DBLP:series/ncs/EibenS15,DeJong2006} are bio-inspired meta-heuristics that can be effectively used to find nearly optimal solutions for optimization problems. Usually an EA starts by generating a set of random solutions, called \emph{population}, following a user-defined description. Then, it evaluates each candidate solution, called \emph{individual}, assigning it a \emph{fitness} value, that describes how good the individual is, with regards to the target problem. New solutions are then generated by the application of \emph{operators} that either mutate a single solution or recombine different existing solutions. After each iteration, called \emph{generation}, the least fit individuals are removed, and the process continues until a user-defined stop condition is met.

Multi-objective optimization problems (MOP) are those where several objectives have to be simultaneously optimized \citep{Mora13paretobased}. Solving a MOP implies maximize or minimize a function composed of several cost functions (one per objective). In these problems the aim is to obtain a set of solutions that are better than the rest, considering all the objectives; this set is known as the Pareto Front. The solutions in this set are {\em non-dominated}, that is, none is better than the others for all the objectives. MOEAs are sub-set of EAs used to solve MOPs. One of the most well-known and referenced MOEAs \citep{Dorronsoro13superlinear} is the {\em Non-dominated Sorting Genetic Algorithm} (NSGA-II) \citep{Deb00NSGAII}. Classical genetic operators are applied to all the individuals, and divided into different ranks (based on the dominance) to be selected for the next generations.

Since the early 2000s, the distributed and parallel Evolutionary Algorithms have been used, mostly to take the most from systems such as clusters or grids \citep{Talbi08Parallel}. But in the case of MOEAs, the distribution and parallelization is not as direct as in single-objective EAs. This is because in different steps of the algorithm a whole set of dependent solutions, the Pareto Front, should be managed as a whole, spending time in gathering all individuals from the different processors or islands.

@@

% ¿Se podría mencionar por algún lado el trabajo en algoritmos de
% hormigas paralelos de Antonio? - JJ Fergu: DONE
In our previous work \citep{Garcia16hpmoon} we used some of
the aforementioned ideas to compare two co-evolutionary dMOEAs. The first
one divided the chromosome in $P$ sections, being $P$ the number of
islands. Every island $p$ only performed the mutation and crossover in
that part ($p_{th}$) of the chromosome, while the fitness was
calculated using the whole individual. After a certain number of
generations, individuals were migrated randomly to other islands. The
performance metrics were calculated at the end of the run. The second
method, coevolution with overlapping islands, used the $p_{th-1}$ and
$p_{th+1}$ sections of each chromosome, besides the $p_{th}$
part. During the same amount of time both methods obtained better
results than a baseline algorithm that dealt with the whole chromosome
in each island for crossover and mutation. We discovered that the
performance using one or another method depends on the number of
sections of the individuals and the number of islands used. This
motivates us to find a new adaptive method to select this number of
sections of the chromosome to use, depending on the number of
islands. Besides, previous experiments were performed on a single
processor synchronous island model with a limited number of islands
(8, 32 and 128). In this paper,  8, 16, 32, 64 and 128 islands have
been used, and this time, the experiments have been run on a real
cluster. Therefore, at the same time we are proposing an adaptive
method to divide the individual search space, we are validating the
previous methods.

@@

In this paper we analyse several coevolutionary algorithms we have implemented by using a parallel multi-objective baseline algorithm (B) as a framework. Our proposed methods are based in NSGA-II, as almost all the previous works discussed before \cite{Dorronsoro13superlinear,Durillo08masterslave,Hiroyasu07discussion,Deb03distributed,Xiao03specialized,Wang09parallel,Martens13asynchronous}. Therefore, we have used a basic distributed NSGA-II algorithm without overlapping sections as the baseline (B). The population is distributed among $P$ islands. After a fixed number of generations, one individual in a given island is migrated to another random island, thus avoiding to synchronize the global Pareto Front (PF) every certain number of generations as other methods described in Section 2 do. At the end of the run, the Pareto Fronts of all islands are aggregated into a new one, and the quality measures are evaluated.
Different coevolutionary alternatives can be devised to evolve the subpopulations according to the decision space to be explored by each islands. Here we will consider three methods in which each island only performs crossover and mutation in specific sections of the individuals.

@@

The previously described four approaches have been run with two different chromosome lengths ($L$): 512 and 2048. Different number of islands ($P$) have also been compared: 8, 16, 32, 64 and 128. This maximum number of islands have also been  used in previous work in the literature \citep{Martens13asynchronous}. The crossover and mutation chosen, SBX and polynomial, have also been  used previously by other authors in \citep{Durillo08masterslave}.

>>>>

%This is not state of the art. If it's not an EC journal, it should go
%to the introduction - JJ
Evolutionary algorithms (EAs) \citep{DBLP:series/ncs/EibenS15,DeJong2006} are bio-inspired meta-heuristics that can be effectively used to find nearly optimal solutions for optimization problems. Usually an EA starts by generating a set of random solutions, called \emph{population}, following a user-defined description. Then, it evaluates each candidate solution, called \emph{individual}, assigning it a \emph{fitness} value, that describes how good the individual is, with regards to the target problem. New solutions are then generated by the application of \emph{operators} that either mutate a single solution or recombine different existing solutions. After each iteration, called \emph{generation}, the least fit individuals are removed, and the process continues until a user-defined stop condition is met.

% Not EA either. It's an explanation of the techniques you are using,
% thus Intro - JJ
Multi-objective optimization problems (MOP) are those where several
objectives have to be simultaneously optimized
\citep{Mora13paretobased}. Solving a MOP implies maximize or minimize
a function composed of several cost functions (one per objective). In
these problems the aim is to obtain a set of solutions that are better
than the rest, considering all the objectives; this set is known as
the Pareto Front. The solutions in this set are {\em non-dominated},
that is, none is better than the others for all the objectives. MOEAs
are sub-set of EAs used to solve MOPs. One of the most well-known and
referenced MOEAs \citep{Dorronsoro13superlinear} is the {\em
  Non-dominated Sorting Genetic Algorithm} (NSGA-II)
\citep{Deb00NSGAII}. Classical genetic operators are applied to all
the individuals, and divided into different ranks (based on the
dominance) to be selected for the next generations.

Since the early 2000s distributed and parallel Evolutionary
Algorithms have been used mostly to leverage systems such
as clusters or grids \citep{Talbi08Parallel}, since a first-order
approach to EA parallelization is as easy as dividing the population
and add a simple communication mechanism among them. But in the case of
MOEAs, the distribution and parallelization is not as direct as in
single-objective EAs. This is because in different steps of the
algorithm a whole set of dependent solutions, the Pareto Front, should
be managed as a whole, spending time in gathering all individuals from
the different processors or islands.

@@

In our previous work \citep{Garcia16hpmoon} we used some of
the aforementioned ideas to compare two co-evolutionary dMOEAs. The first
one divided the chromosome in $P$ sections, being $P$ the number of
islands. Every island $p$ only performed the mutation and crossover in
that part ($p_{th}$) of the chromosome, while the fitness was
calculated using the whole individual. After a certain number of
generations, individuals were migrated randomly to other islands. The
performance metrics were calculated at the end of the run. The second
method, coevolution with overlapping islands, used the $p_{th-1}$ and
$p_{th+1}$ sections of each chromosome, besides the $p_{th}$
part. During the same amount of time both methods obtained better
results than a baseline algorithm that dealt with the whole chromosome
in each island for crossover and mutation. We discovered that the
performance using one or another method depends on the number of
sections of the individuals and the number of islands used. This
motivates us to find a new adaptive method to select this number of
sections of the chromosome to use, depending on the number of
islands. Besides, previous experiments were performed on a single
processor synchronous island model with a limited number of islands
(8, 32 and 128). In this paper,  8, 16, 32, 64 and 128 islands have
been used, and this time, the experiments have been run on a real
cluster. Therefore, at the same time we are proposing an adaptive
method to divide the individual search space, we are validating the
previous methods.

@@

In this paper we analyse several coevolutionary algorithms we have
implemented by using a parallel multi-objective baseline algorithm (B)
as a basis for comparisons. Our proposed methods %methods or method?
                                %Which methods? - JJ
are based in NSGA-II, as almost all the previous works discussed
before
\cite{Dorronsoro13superlinear,Durillo08masterslave,Hiroyasu07discussion,Deb03distributed,Xiao03specialized,Wang09parallel,Martens13asynchronous}. % This is SoA stuff - JJ
Therefore, we have used a basic distributed NSGA-II algorithm without
overlapping sections as the baseline (B).

%This probably should deserve a "Algorithm" environment too - JJ
This basic distributed algorithms spreads the population among $P$
islands; after a fixed number of generations,
one individual in a given island is migrated to another random island,
thus avoiding to synchronize the global Pareto Front (PF) every
certain number of generations as other methods described in Section 2
do. At the end of the run, the Pareto Fronts of all islands are
aggregated into a new one, and the quality measures are evaluated. 
Different coevolutionary alternatives can be devised to evolve the
subpopulations according to the decision space to be explored by each
islands. Here we will consider three methods in which each island only
performs crossover and mutation in specific sections of the
individuals.

@@

The previously described four approaches have been run with two
different chromosome lengths ($L$): 512 and 2048. Different number of
islands ($P$) have also been compared: 8, 16, 32, 64 and 128. This
maximum number of islands have also been  used in previous work in the
literature \citep{Martens13asynchronous}. The crossover and mutation
chosen, SBX and polynomial, have also been  used previously by other
authors in \citep{Durillo08masterslave}.  % SBX is first mentioned
                                % here. Has to explain - JJ
